### 2022-07-09 notes/theory.tex
theory notes

\documentclass[11pt]{article}
\RequirePackage{fullpage}
\RequirePackage[font=small,labelfont=bf]{caption}
\RequirePackage{amsmath,amssymb,amsthm}
\RequirePackage{graphicx}
\RequirePackage[hidelinks]{hyperref}
\RequirePackage{subcaption}
\RequirePackage{wasysym}
\RequirePackage{authblk}
\RequirePackage{bm}
\RequirePackage{bbm}
% \RequirePackage[osf]{mathpazo}
\RequirePackage[bibstyle=authoryear,citestyle=authoryear-comp,maxbibnames=9,maxcitenames=1,backend=biber,natbib=true,uniquelist=false,hyperref=true]{biblatex}
\usepackage{color}
\usepackage{nicefrac}

\renewcommand{\P}{\mathbb{P}}
\newcommand{\E}{\mathbb{E}}
\newcommand{\V}{\text{V}}
\DeclareMathOperator{\var}{Var}
\DeclareMathOperator{\cov}{cov}

\addbibresource{biblio.bib}

\title{Statistical Methods}

\begin{document}
\maketitle

\section*{Likelihood}

The likelihood in Eyalshiv et al. (2016) has the form,

\begin{align}
  \log\mathcal{L} = \sum_{v \in \mathcal{V}} \sum_{i \ne j \in \mathcal{S}} \log(P(O_{i,j}(v) | \theta))
\end{align}

where $\mathcal{V}$ is the set of putatively neutral sites, $\mathcal{S}$ is
the set of samples, and $\theta$ are the BGS parameters. The indicator variable
$O_{i,j}(v)$ is 1 if samples $i$ and $j$ are different at site $v$, and zero
otherwise. Thus as they specify in the paper,

\begin{align}
  P(O_{i,j}(v) | \theta) = 
    \begin{cases}
      \pi(v | \theta), & O_{i,j}(v) = 1 \\
      1-\pi(v | \theta), & O_{i,j}(v) = 0 \\
    \end{cases}.
\end{align}

The total size of the set of samples $\mathcal{S}$ is $n_\mathcal{S} =
|\mathcal{S}|$. Assuming all sites are biallelic, we can simplify the
calculation of the likelihood by noting that for a site $v$'s vector of allele
counts $[c_1, c_2]$, the total number of pairwise combinations with the same
alleles is

\begin{align}
  n_s(v) = {c_1 \choose 2} + {c_2 \choose 2}
\end{align}

and the number of different pairwise combinations is

\begin{align}
  n_d(v) = n_T - n_s(v)
\end{align}

where $n_T = n_\mathcal{S} (n_\mathcal{S} - 1) / 2$  is the total number of
pairwise combinations across the sample set $\mathcal{S}$. Furthermore, note
that we can partition the set $\mathcal{V}$ of neutral sites into polymorphic
($\mathcal{P}$) and fixed sites ($\mathcal{F}$), i.e. $\mathcal{V} =
\mathcal{P} \cup \mathcal{F}$ and $\mathcal{P} \cap \mathcal{F} = \varnothing$.
For all $v \in \mathcal{F}$, $n_d(v) = 0$ and $n_s(v) = n_T$.

Then,

\begin{align}
  \log\mathcal{L} &= \sum_{v \in \mathcal{V}} \left[\log(\pi(v | \theta)) n_\text{D}(v) + \log(1-\pi(v | \theta)) n_\text{S}(v)\right] \\
                  &= \sum_{v \in \mathcal{P}} \left[\log(\pi(v | \theta)) n_\text{D}(v) + \log(1-\pi(v | \theta)) n_\text{S}(v)\right] + \sum_{v \in \mathcal{F}} \log(1-\pi(v | \theta)) n_T  \\
\end{align}

\section{B Scores}

\end{document}